---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# AutoCDAScorer


<!-- badges: start -->
[![codecov](https://codecov.io/gh/joshuandwilliams/AutoCDAScorer/graph/badge.svg?token=DVSFFFKKQ4)](https://codecov.io/gh/joshuandwilliams/AutoCDAScorer)
[![R CMD check](https://github.com/joshuandwilliams/AutoCDAScorer/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/joshuandwilliams/AutoCDAScorer/actions/workflows/R-CMD-check.yaml)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
![R](https://img.shields.io/badge/R-%3E=4.4.2-1e90ff?logo=r)
[![tensorflow version](https://img.shields.io/badge/tensorflow-v2.16.2-orange)](https://www.tensorflow.org/)
<!-- badges: end -->

The goal of AutoCDAScorer is to make scoring cell death areas on UV spectra agroinfiltration images faster, more consistent, and less subjective.

AutoCDAScorer uses deep learning models trained on UV spectra agroinfiltration images to automatically predict cell death severity between 0 (no cell death) and 6 (strong cell death).

## Installation

You can install the development version of AutoCDAScorer from [GitHub](https://github.com/) with:
      
```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("joshuandwilliams/AutoCDAScorer")
```

The `AutoCDAScorer` package uses TensorFlow. If you already have TensorFlow installed you can use `reticulate::use_condaenv()` to specify the correct environment for R to use. Otherwise you will need to install TensorFlow. You can do this from within R:

```r
library(AutoCDAScorer)
install_tensorflow()
```

This installation only needs to be done once. After that, TensorFlow will remain installed unless you actively remove it.

## Usage

The package vignette provides a workflow for loading images, assessing their suitability, and predicting their score.

```{r, eval = FALSE}
browseVignettes("AutoCDAScorer")
```
