---
title: "dev_test_pipeline"
output: html_document
---

A test script to run the AutoCDAScorer functions on example data.


# Load libraries
```{r}
library(AutoCDAScorer)
library(reticulate)
use_condaenv("/Users/jowillia/opt/anaconda3/envs/r-tensorflow")
```


# Load model and images. View example image
```{r}
# Load model
model <- load_cda_model("base_cnn")

# Load original images
original_dataset <- load_images_and_labels("/Users/jowillia/Documents/GitHub/Bulletproof_Models/data/cropped_images", image_size=64, labels=TRUE)

# Load new images
new_dataset <- load_images_and_labels("/Users/jowillia/Documents/GitHub/Bulletproof_Models/data/bio_test", image_size=64, labels=TRUE)
```
```{r}
show_test_image(original_dataset)
show_test_image(new_dataset)
```

# Build PCA from original images and transform new images using generated PCs
```{r}
result_pca <- run_pca_python(original_dataset, n_components=5)
transformed_new_features <- pca_transform_python(new_dataset, result_pca)
```

```{r}
print(transformed_new_features)
```

# Diagnostic plots
Replace interactive plot with all-against-all plots. Here are three types: target plots, contour plots, and convex hull plots.
```{r}
target_plot <- diagnostic_pca_all_against_all(
  result_pca,
  transformed_new_features,
  num_pcs = 5,
  plot_type = "target",
  num_ellipses = 3
)
target_plot

contour_plot <- diagnostic_pca_all_against_all(
  result_pca,
  transformed_new_features,
  num_pcs = 5,
  plot_type = "density",
  num_bins = 5
)
contour_plot

convexhull_plot <- diagnostic_pca_all_against_all(
  result_pca,
  transformed_new_features,
  num_pcs = 5,
  plot_type = "convexhull"
)
convexhull_plot
```

# Make score predictions
```{r}
new_images_bgr <- rgb_to_bgr(new_dataset)
new_predictions <- predict_score("base_cnn", new_images_bgr, softmax = FALSE)
table(new_predictions)
```
