---
title: "dev_test_pipeline"
output: html_document
---

A test script to run the AutoCDAScorer functions on example data.


# Load libraries
```{r}
library(AutoCDAScorer)
library(reticulate)
use_condaenv("/Users/jowillia/opt/anaconda3/envs/r-tensorflow")
```


# Load model and images. View example image
```{r}
# Load original images
original_dataset <- load_images("/Users/jowillia/Documents/GitHub/Bulletproof_Models/data/cropped_images")
show_test_image(original_dataset)

# Load new images
new_dataset <- load_images("/Users/jowillia/Documents/GitHub/Bulletproof_Models/data/bio_test")
show_test_image(new_dataset)
```

# Build PCA from original images and transform new images using generated PCs
```{r}
run_pca_python <- function(data, n_components = 10) {
  # Check Python installation and environment
  if (!reticulate::py_available(initialize = TRUE)) {
    stop("Error: Python is not available. Please install Python and configure reticulate.")
  }
  if (!reticulate::py_module_available("numpy")) {
    stop("Error: The 'numpy' module is not installed in the Python environment. Install it using 'pip install numpy'.")
  }
  if (!reticulate::py_module_available("sklearn.decomposition")) {
    stop("Error: The 'scikit-learn' module is not installed in the Python environment. Install it using 'pip install scikit-learn'.")
  }
  
  # Import Python modules
  np <- reticulate::import("numpy")
  sklearn_decomposition <- reticulate::import("sklearn.decomposition")

  images <- data$images
  images_array <- np$array(images)
  images_array <- np$reshape(images_array, c(dim(images_array)[1], as.integer(-1)))
  
  # Validate n_components
  if (n_components > dim(images_array)[2]) {
    stop("Error: 'n_components' cannot be greater than the number of features in the input data.")
  }

  # Perform PCA
  pca <- sklearn_decomposition$PCA(n_components=as.integer(floor(n_components)))
  features_pca <- pca$fit_transform(images_array)

  # Convert Python objects to R format
  features_pca_r <- reticulate::py_to_r(features_pca)
  explained_variance_r <- reticulate::py_to_r(pca$explained_variance_ratio_)
  principal_components_r <- reticulate::py_to_r(pca$components_)
  center_r <- reticulate::py_to_r(pca$mean_)  # Equivalent to prcomp$center in R

  # Return only the essential PCA components
  return(list(
    features_pca = features_pca_r,
    explained_variance = explained_variance_r[1:n_components],
    principal_components = principal_components_r[1:n_components, ],
    center = center_r
  ))
}

result_pca <- run_pca_python(original_dataset, n_components=5)
#saveRDS(result_pca, "~/Packages/AutoCDAScorer/inst/extdata/base_cnn_pca.rds")
```

# Diagnostic plots
Replace interactive plot with all-against-all plots. Here are three types: target plots, contour plots, and convex hull plots.
```{r}
target_plot <- diagnostic_pca(
  "base_cnn",
  new_dataset,
  num_pcs = 5,
  plot_type = "target",
  num_ellipses = 3
)
target_plot

contour_plot <- diagnostic_pca(
  "base_cnn",
  new_dataset,
  num_pcs = 5,
  plot_type = "density",
  num_bins = 5
)
contour_plot

convexhull_plot <- diagnostic_pca(
  "base_cnn",
  new_dataset,
  num_pcs = 5,
  plot_type = "convexhull"
)
convexhull_plot
```

Good and bad examples
```{r}
indices <- grep("^DSC_0105", original_dataset$filenames)
good_dataset <- list(images = original_dataset$images[indices, , ,]) # Subset of training data

convexhull_plot <- diagnostic_pca(
  "base_cnn",
  good_dataset,
  num_pcs = 5,
  plot_type = "convexhull"
)
convexhull_plot
```

```{r}
saveRDS(good_dataset, file = "~/Packages/AutoCDAScorer/inst/extdata/example_dataset/training_subset/good_dataset.rds")
```

```{r}
bad_dataset <- good_dataset
bad_dataset$images <- bad_dataset$images[,,,c(3,1,2)] # Mix up the colour channels

convexhull_plot <- diagnostic_pca(
  "base_cnn",
  bad_dataset,
  num_pcs = 5,
  plot_type = "convexhull"
)
convexhull_plot
```

# Make score predictions
```{r}
new_dataset_bgr <- rgb_to_bgr(new_dataset)
predictions <- predict_score("base_cnn", new_dataset_bgr, softmax = FALSE)
table(predictions)
```

# Save predictions
```{r}
annotations_to_csv(new_dataset_bgr, predictions, "~/Packages/AutoCDAScorer/dev/annotations.csv")
```
