---
title: "dev_test_pipeline"
output: html_document
---

A test script to run the AutoCDAScorer functions on example data.


# Load libraries
```{r}
library(AutoCDAScorer)
library(reticulate)
use_condaenv("/Users/jowillia/opt/anaconda3/envs/r-tensorflow")
```


# Load model and images. View example image
```{r}
# Load model
model <- load_cda_model("base_cnn")

# Load original images
original_dataset <- load_images_and_labels("/Users/jowillia/Documents/GitHub/Bulletproof_Models/data/cropped_images", image_size=64, labels=TRUE)

# Load new images
new_dataset <- load_images_and_labels("/Users/jowillia/Documents/GitHub/Bulletproof_Models/data/bio_test", image_size=64, labels=TRUE)
```
```{r}
show_test_image(original_dataset)
show_test_image(new_dataset)
```

# Build PCA from original images and transform new images using generated PCs
```{r}
result_pca <- run_pca_python(original_dataset, n_components=5)
transformed_new_features <- pca_transform_python(new_dataset, result_pca)
```

```{r}
print(transformed_new_features)
```

# Diagnostic plots
Replace interactive plot with all-against-all plots. Here are three types: target plots, contour plots, and convex hull plots.
```{r}
#' Diagnostic PCA Scatter Plots for All Principal Component Pairs
#'
#' This function generates a grid of scatter plots comparing original and new features across
#' all pairwise combinations of principal components (PCs). It supports different visualization
#' styles: target-based ellipses, density contours, or convex hulls.
#'
#' @param pca A list containing the PCA model output, including `original_features` and `explained_variance`.
#' @param new_features A matrix of new features transformed by PCA.
#' @param num_pcs An integer specifying the number of principal components to consider.
#' @param plot_type A string specifying the type of plot: `"target"`, `"density"`, or `"convexhull"`.
#' @param num_ellipses An integer specifying the number of ellipses for the `"target"` plot type (default: 3).
#' @param num_bins An integer specifying the number of bins for the `"density"` plot type (default: 3).
#'
#' @return A combined ggplot2 object displaying all pairwise PCA scatter plots in a grid format.
#'
#' @import ggplot2
#' @importFrom cowplot ggdraw draw_plot draw_label plot_grid
#' @export
diagnostic_pca_all_against_all <- function(pca, new_features, num_pcs, plot_type, num_ellipses = 3, num_bins = 3) {
  original_features <- pca$features_pca
  explained_variance <- pca$explained_variance

  # Input validation
  if (length(original_features) == 0 || length(new_features) == 0) {
    stop("Error: Input matrices cannot be empty.")
  }
  if (!is.matrix(original_features)) {
    stop("Error: 'original_features' must be a matrix.")
  }
  if (!is.matrix(new_features)) {
    stop("Error: 'new_features' must be a matrix.")
  }
  if (ncol(original_features) != ncol(new_features)) {
    stop("Original and new features must have the same number of principal components.")
  }
  if (num_pcs > ncol(original_features)) {
    stop("num_pcs exceeds the number of available principal components.")
  }

  # Generate all possible PC combinations
  generate_pc_combinations <- function(num_pcs) {
    combinations <- expand.grid(PC_a = seq_len(num_pcs), PC_b = seq_len(num_pcs))
    combinations <- combinations[combinations$PC_a < combinations$PC_b, ]
    return(combinations)
  }

  # Generate combinations
  plot_combinations <- generate_pc_combinations(num_pcs)

  # Create list of plots
  # Lookup table to call the appropriate plotting function based on plot_type
  plot_list <- lapply(seq_len(nrow(plot_combinations)), function(i) {
    PC_a <- plot_combinations[i, "PC_a"]
    PC_b <- plot_combinations[i, "PC_b"]

    if (plot_type == "target") {
      return(pca_plot_with_target(original_features, new_features, explained_variance, PC_a, PC_b, num_pcs, num_ellipses))
    } else if (plot_type == "density") {
      return(pca_plot_with_density(original_features, new_features, explained_variance, PC_a, PC_b, num_pcs, num_bins))
    } else if (plot_type == "convexhull") {
      return(pca_plot_with_convex_hull(original_features, new_features, PC_a, PC_b))
    } else {
      stop("Invalid plot type.")
    }
  })

  create_full_plot_grid <- function(plot_list, num_pcs, explained_variance, plot_combinations) {
    full_grid <- matrix(list(NULL), nrow = num_pcs, ncol = num_pcs)
  
    for (i in seq_len(nrow(plot_combinations))) {
      pc_a <- plot_combinations[i, "PC_a"]
      pc_b <- plot_combinations[i, "PC_b"]
      
      # Insert the scatter plot in the original position
      full_grid[[pc_a, pc_b]] <- plot_list[[i]]
      
      # Compute explained variance sum
      total_variance <- explained_variance[pc_a] + explained_variance[pc_b]

      # Create a circle plot in the mirrored position
      circle_plot <- ggplot(data.frame(x = 1, y = 1, size = total_variance, label = paste0(round(total_variance * 100, 1), "%"))) +
        geom_point(aes(x = x, y = y, size = size), shape = 21, fill = "blue", color = "black") +
        scale_size_continuous(
          range = c(1, 20),  # Adjust the range for circle sizes
          limits = c(0, 1)   # Ensure variance scales from 0 to 1
        ) +
        geom_text(
          aes(x = x, y = y + 0.75, label = label),
          size = 4,  # Adjust text size
          color = "black"  # Set text color for visibility
        ) +
        theme_void() + theme(legend.position = "none") + 
        coord_cartesian(clip = "off") + 
        ylim(0, 1.75)
            
      # Insert circle in mirrored position
      full_grid[[pc_b, pc_a]] <- circle_plot
    }
  
    return(full_grid)
  }

  # Create full plot grid
  full_plot_grid <- create_full_plot_grid(plot_list, num_pcs, explained_variance, plot_combinations)

  # Create combined plot
  combined_plot <- ggdraw() +
    draw_plot(
      plot_grid(
        plotlist = as.list(full_plot_grid[, num_pcs:1]),
        ncol = num_pcs,
        nrow = num_pcs,
        align = 'none',
        axis = 'tblr'
      ),
      x = 0.05,     # Move plot right
      y = 0.05,     # Move plot up
      width = 0.9, # Adjust plot width
      height = 0.9 # Adjust plot height
    ) +
    # Add shared x-axis principal component label
    draw_label(
      "PC",
      x = 0.5,
      y = 0.02,
      size = 10
    ) +
    # Add shared y-axis principal component label
    draw_label(
      "PC",
      x = 0.02,
      y = 0.5,
      angle = 90,
      size = 10
    )

  # Add x-axis tick labels at precise coordinates
  for (i in seq_len(num_pcs)) {
    x_pos <- 0.05 + ((i - 0.5) / num_pcs) * 0.9  # Adjust x position relative to plot width
    combined_plot <- combined_plot +
      draw_label(
        as.character(i),
        x = x_pos,
        y = 0.04,
        size = 10
      )
  }

  # Add y-axis tick labels at precise coordinates
  for (i in seq_len(num_pcs)) {
    y_pos <- 0.05 + ((i - 0.5) / num_pcs) * 0.9  # Adjust y position relative to plot height
    combined_plot <- combined_plot +
      draw_label(
        as.character(i),
        x = 0.04,
        y = y_pos,
        angle = 90,
        size = 10
      )
  }
  # Add legend
  combined_plot <- combined_plot +
    draw_plot(
      ggplot() +
        geom_point(
          aes(x = 1, y = 0.5, color = "Original"),
          size = 3, alpha = 0
        ) +
        geom_point(
          aes(x = 2, y = 0.5, color = "New"),
          size = 3, alpha = 0
        ) +
        scale_color_manual(
          values = c("Original" = "grey", "New" = "red")
        ) +
        theme_void() +
        theme(
          legend.position = "bottom",
          legend.title = element_blank()
        ) +
        guides(
          color = guide_legend(override.aes = list(size = 5, alpha = 1))
        ),
      x = 0.4,
      y = 0.95,
      width = 0.2,
      height = 0.05
    )

  return(combined_plot)
}
```

```{r}
target_plot <- diagnostic_pca_all_against_all(
  result_pca,
  transformed_new_features,
  num_pcs = 5,
  plot_type = "target",
  num_ellipses = 3
)
target_plot

contour_plot <- diagnostic_pca_all_against_all(
  result_pca,
  transformed_new_features,
  num_pcs = 5,
  plot_type = "density",
  num_bins = 5
)
contour_plot

convexhull_plot <- diagnostic_pca_all_against_all(
  result_pca,
  transformed_new_features,
  num_pcs = 5,
  plot_type = "convexhull"
)
convexhull_plot
```

# Make score predictions
```{r}
new_images_bgr <- rgb_to_bgr(new_dataset)
new_predictions <- predict_score("base_cnn", new_images_bgr, softmax = FALSE)
table(new_predictions)
```
